---
title: "Type Conversion Model"
author: "Anne-Marie Parkinson"
date: "February 2, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# load packages
```{r}
library(tidyverse)
library(here)
library(dplyr)
library(pROC)
library(ROCR)
```

# add data
```{r}
sf2 <-read_csv(here::here("Shade Fraction Data Sheets and Code", "Final Zaca MCF SF_MainData.csv")) %>% 
  dplyr::mutate(mortality_qualitative= fct_relevel(mortality_qualitative, levels=c("None", "Low", "Moderate", "High"))) %>% 
  dplyr::mutate(other_mortality = fct_relevel(other_mortality, levels = c("None", "Low", "Moderate", "High", "Type Converted")))

gv2 <- read_csv(here::here("Shade Fraction Data Sheets and Code", "Zaca MCF variables for model_plus GV.csv")) %>% 
  dplyr::mutate(mortality_qual= fct_relevel(mortality_qual, levels=c("None", "Low", "Moderate", "High"))) %>% 
  rename(mortality_qualitative = mortality_qual) 

```

# sample size 
```{r}
count <- sf2 %>% 
  group_by(new_model_class) %>% 
  count()
```

# type conversion model trials

## shade fractions: one response variable

```{r}
# make lists to later store values
acc_list <- list()
table_values <- list()
model_aic <-list()
auc_list <- list ()
#prediction_list <- list()

#type conversion model ------------------------------------
for (i in 1:500)
  {
  #divide the response variable into equal number of observations
  tc_yes <- sf2[which(sf2$new_model_class == "1"),]
  tc_no <- sf2[which(sf2$new_model_class == "0"),]
  
  # pull 175 random observations from the high class and separate
  yes <- sample(1:nrow(tc_yes), 99 , replace = FALSE)
  polygon_data_high_train <- tc_yes[yes,]
  # pull 175 random observations from the low class and separate
  no <- sample(1:nrow(tc_no), 99 , replace = FALSE)
  polygon_data_low_train <- tc_no[no,]
  
  # combine all random observations
  train <- do.call("rbind", list(polygon_data_high_train, polygon_data_low_train))
  
  tc <- glm(new_model_class~shade_2018, data=train, family="binomial")
  sum <-summary(tc)
  sum
  aic<- sum$aic
  aic
  model_aic <- append(model_aic, aic)
  
  # predict data 
  pred<-predict(tc, train, type="response") 
  
  # error rate for data
  pred1 <- ifelse (pred>0.5,1,0)
  tab<- table(Predicted=pred1, Actual=train$new_model_class)
  write.table(tab, file = "type conversion train response variable accuracy.csv", sep=",",col.names=NA, append=T)

  acc <-1-sum(diag(tab))/sum(tab)   #error calculation. gives % observations predicted incorrectly 
  acc <- 1- acc 
  acc_list <- append(acc_list, acc)
  
  #auc score
  pred <- predict(tc,train)
  pred1 <-prediction(pred, train$new_model_class)
  roc <- performance(pred1, "tpr", "fpr")
  auc <-performance(pred1, "auc")
  auc1 <- unlist(slot(auc, "y.values"))
  auc_list <- append(auc_list, auc1)
  
}


model_aic <- as.data.frame(unlist(model_aic))
acc_list <- as.data.frame(unlist(acc_list))
auc_list <-as.data.frame(unlist(auc_list))

combo <- cbind(model_aic, acc_list, auc_list)
write.table(model_aic, file="model_aic.csv", sep=",")

# table of mean model outputs ---------------------------------------------------

sf_combo_data <- read_csv(here::here("Shade Fraction Data Sheets and Code", "model_aic.csv"))
# the column headers are off, they 

# read in data and make it R friendly 
type_converted <- read_csv(here::here("Shade Fraction Data Sheets and Code", "type conversion train response variable accuracy.csv")) %>% 
  rename(type_converted = '0')
not_type_converted <- read_csv(here::here("Shade Fraction Data Sheets and Code", "type conversion train response variable accuracy.csv")) %>% rename(not_type_converted = '1')

type_converted <- type_converted %>% 
  select (X1, type_converted) %>% 
  drop_na() %>% 
  filter (X1 == 0) %>% 
  mutate(class_acc = type_converted/99)
type_converted_accuracy <- type_converted %>% 
   summarise(mean = mean(class_acc))

not_type_converted <- not_type_converted %>% 
  select (X1, not_type_converted) %>% 
  drop_na() %>% 
  filter (X1 == 1) %>% 
  mutate(class_acc = not_type_converted/99)
not_type_converted_accuracy <- not_type_converted %>% 
   summarise(mean = mean(class_acc))
```

## Shade fractions: two response variables
```{r}

# make lists to later store values
acc_list <- list()
table_values <- list()
model_aic <-list()
auc_list <- list ()
#prediction_list <- list()

#type conversion model ------------------------------------
for (i in 1:500)
  {
  #divide the response variable into equal number of observations
  tc_yes <- sf2[which(sf2$new_model_class == "1"),]
  tc_no <- sf2[which(sf2$new_model_class == "0"),]
  
  # pull 175 random observations from the high class and separate
  yes <- sample(1:nrow(tc_yes), 99 , replace = FALSE)
  polygon_data_high_train <- tc_yes[yes,]
  # pull 175 random observations from the low class and separate
  no <- sample(1:nrow(tc_no), 99 , replace = FALSE)
  polygon_data_low_train <- tc_no[no,]
  
  # combine all random observations
  train <- do.call("rbind", list(polygon_data_high_train, polygon_data_low_train))
  
  tc <- glm(new_model_class~shade_2018+max_2000_2006, data=train, family="binomial")
  sum <-summary(tc)
  sum
  aic<- sum$aic
  aic
  model_aic <- append(model_aic, aic)
  
  # predict data 
  pred<-predict(tc, train, type="response") 
  
  # error rate for data
  pred1 <- ifelse (pred>0.5,1,0)
  tab<- table(Predicted=pred1, Actual=train$new_model_class)
  write.table(tab, file = "type conversion train response variable accuracy.csv", sep=",",col.names=NA, append=T)

  acc <-1-sum(diag(tab))/sum(tab)   #error calculation. gives % observations predicted incorrectly 
  acc <- 1- acc 
  acc_list <- append(acc_list, acc)
  
  #auc score
  pred <- predict(tc,train)
  pred1 <-prediction(pred, train$new_model_class)
  roc <- performance(pred1, "tpr", "fpr")
  auc <-performance(pred1, "auc")
  auc1 <- unlist(slot(auc, "y.values"))
  auc_list <- append(auc_list, auc1)
  
}


model_aic <- as.data.frame(unlist(model_aic))
acc_list <- as.data.frame(unlist(acc_list))
auc_list <-as.data.frame(unlist(auc_list))

combo <- cbind(model_aic, acc_list, auc_list)
write.table(model_aic, file="model_aic.csv", sep=",")

# table of mean model outputs ---------------------------------------------------

sf_combo_data <- read_csv(here::here("Shade Fraction Data Sheets and Code", "model_aic.csv"))
# the column headers are off, they 

# read in data and make it R friendly 
type_converted <- read_csv(here::here("Shade Fraction Data Sheets and Code", "type conversion train response variable accuracy.csv")) %>% 
  rename(type_converted = '0')
not_type_converted <- read_csv(here::here("Shade Fraction Data Sheets and Code", "type conversion train response variable accuracy.csv")) %>% rename(not_type_converted = '1')

type_converted <- type_converted %>% 
  select (X1, type_converted) %>% 
  drop_na() %>% 
  filter (X1 == 0) %>% 
  mutate(class_acc = type_converted/99)
type_converted_accuracy <- type_converted %>% 
   summarise(mean = mean(class_acc))

not_type_converted <- not_type_converted %>% 
  select (X1, not_type_converted) %>% 
  drop_na() %>% 
  filter (X1 == 1) %>% 
  mutate(class_acc = not_type_converted/99)
not_type_converted_accuracy <- not_type_converted %>% 
   summarise(mean = mean(class_acc))


```


## shade fractions: three response variables
```{r}
# make lists to later store values
acc_list <- list()
table_values <- list()
model_aic <-list()
auc_list <- list ()
#prediction_list <- list()

#type conversion model ------------------------------------
for (i in 1:500)
  {
  #divide the response variable into equal number of observations
  tc_yes <- sf2[which(sf2$new_model_class == "1"),]
  tc_no <- sf2[which(sf2$new_model_class == "0"),]
  
  # pull 175 random observations from the high class and separate
  yes <- sample(1:nrow(tc_yes), 99 , replace = FALSE)
  polygon_data_high_train <- tc_yes[yes,]
  # pull 175 random observations from the low class and separate
  no <- sample(1:nrow(tc_no), 99 , replace = FALSE)
  polygon_data_low_train <- tc_no[no,]
  
  # combine all random observations
  train <- do.call("rbind", list(polygon_data_high_train, polygon_data_low_train))
  
  tc <- glm(new_model_class~shade_2018+max_2000_2006+min_2000_200, data=train, family="binomial")
  sum <-summary(tc)
  sum
  aic<- sum$aic
  aic
  model_aic <- append(model_aic, aic)
  
  # predict data 
  pred<-predict(tc, train, type="response") 
  
  # error rate for data
  pred1 <- ifelse (pred>0.5,1,0)
  tab<- table(Predicted=pred1, Actual=train$new_model_class)
  write.table(tab, file = "type conversion train response variable accuracy.csv", sep=",",col.names=NA, append=T)

  acc <-1-sum(diag(tab))/sum(tab)   #error calculation. gives % observations predicted incorrectly 
  acc <- 1- acc 
  acc_list <- append(acc_list, acc)
  
  #auc score
  pred <- predict(tc,train)
  pred1 <-prediction(pred, train$new_model_class)
  roc <- performance(pred1, "tpr", "fpr")
  auc <-performance(pred1, "auc")
  auc1 <- unlist(slot(auc, "y.values"))
  auc_list <- append(auc_list, auc1)
  
}


model_aic <- as.data.frame(unlist(model_aic))
acc_list <- as.data.frame(unlist(acc_list))
auc_list <-as.data.frame(unlist(auc_list))

combo <- cbind(model_aic, acc_list, auc_list)
write.table(model_aic, file="model_aic.csv", sep=",")

# table of mean model outputs ---------------------------------------------------

sf_combo_data <- read_csv(here::here("Shade Fraction Data Sheets and Code", "model_aic.csv"))
# the column headers are off, they 

# read in data and make it R friendly 
type_converted <- read_csv(here::here("Shade Fraction Data Sheets and Code", "type conversion train response variable accuracy.csv")) %>% 
  rename(type_converted = '0')
not_type_converted <- read_csv(here::here("Shade Fraction Data Sheets and Code", "type conversion train response variable accuracy.csv")) %>% rename(not_type_converted = '1')

type_converted <- type_converted %>% 
  select (X1, type_converted) %>% 
  drop_na() %>% 
  filter (X1 == 0) %>% 
  mutate(class_acc = type_converted/99)
type_converted_accuracy <- type_converted %>% 
   summarise(mean = mean(class_acc))

not_type_converted <- not_type_converted %>% 
  select (X1, not_type_converted) %>% 
  drop_na() %>% 
  filter (X1 == 1) %>% 
  mutate(class_acc = not_type_converted/99)
not_type_converted_accuracy <- not_type_converted %>% 
   summarise(mean = mean(class_acc))
```

## shade fractions: four response variables

```{r}
# make lists to later store values
acc_list <- list()
table_values <- list()
model_aic <-list()
auc_list <- list ()
#prediction_list <- list()

#type conversion model ------------------------------------
for (i in 1:500)
  {
  #divide the response variable into equal number of observations
  tc_yes <- sf2[which(sf2$new_model_class == "1"),]
  tc_no <- sf2[which(sf2$new_model_class == "0"),]
  
  # pull 175 random observations from the high class and separate
  yes <- sample(1:nrow(tc_yes), 99 , replace = FALSE)
  polygon_data_high_train <- tc_yes[yes,]
  # pull 175 random observations from the low class and separate
  no <- sample(1:nrow(tc_no), 99 , replace = FALSE)
  polygon_data_low_train <- tc_no[no,]
  
  # combine all random observations
  train <- do.call("rbind", list(polygon_data_high_train, polygon_data_low_train))
  
  tc <- glm(new_model_class~shade_2018+max_2000_2006+min_2000_200+prefire, data=train, family="binomial")
  sum <-summary(tc)
  sum
  aic<- sum$aic
  aic
  model_aic <- append(model_aic, aic)
  
  # predict data 
  pred<-predict(tc, train, type="response") 
  
  # error rate for data
  pred1 <- ifelse (pred>0.5,1,0)
  tab<- table(Predicted=pred1, Actual=train$new_model_class)
  write.table(tab, file = "type conversion train response variable accuracy.csv", sep=",",col.names=NA, append=T)

  acc <-1-sum(diag(tab))/sum(tab)   #error calculation. gives % observations predicted incorrectly 
  acc <- 1- acc 
  acc_list <- append(acc_list, acc)
  
  #auc score
  pred <- predict(tc,train)
  pred1 <-prediction(pred, train$new_model_class)
  roc <- performance(pred1, "tpr", "fpr")
  auc <-performance(pred1, "auc")
  auc1 <- unlist(slot(auc, "y.values"))
  auc_list <- append(auc_list, auc1)
  
}


model_aic <- as.data.frame(unlist(model_aic))
acc_list <- as.data.frame(unlist(acc_list))
auc_list <-as.data.frame(unlist(auc_list))

combo <- cbind(model_aic, acc_list, auc_list)
write.table(model_aic, file="model_aic.csv", sep=",")

# table of mean model outputs ---------------------------------------------------

sf_combo_data <- read_csv(here::here("Shade Fraction Data Sheets and Code", "model_aic.csv"))
# the column headers are off, they 

# read in data and make it R friendly 
type_converted <- read_csv(here::here("Shade Fraction Data Sheets and Code", "type conversion train response variable accuracy.csv")) %>% 
  rename(type_converted = '0')
not_type_converted <- read_csv(here::here("Shade Fraction Data Sheets and Code", "type conversion train response variable accuracy.csv")) %>% rename(not_type_converted = '1')

type_converted <- type_converted %>% 
  select (X1, type_converted) %>% 
  drop_na() %>% 
  filter (X1 == 0) %>% 
  mutate(class_acc = type_converted/99)
type_converted_accuracy <- type_converted %>% 
   summarise(mean = mean(class_acc))

not_type_converted <- not_type_converted %>% 
  select (X1, not_type_converted) %>% 
  drop_na() %>% 
  filter (X1 == 1) %>% 
  mutate(class_acc = not_type_converted/99)
not_type_converted_accuracy <- not_type_converted %>% 
   summarise(mean = mean(class_acc))
```

## GV fractions: one response variable
```{r}
# make lists to later store values
acc_list <- list()
table_values <- list()
model_aic <-list()
auc_list <- list ()
#prediction_list <- list()

#type conversion model ------------------------------------
for (i in 1:500)
  {
  #divide the response variable into equal number of observations
  tc_yes <- gv2[which(gv2$new_model_class == "1"),]
  tc_no <- gv2[which(gv2$new_model_class == "0"),]
  
  # pull 175 random observations from the high class and separate
  yes <- sample(1:nrow(tc_yes), 99 , replace = FALSE)
  polygon_data_high_train <- tc_yes[yes,]
  # pull 175 random observations from the low class and separate
  no <- sample(1:nrow(tc_no), 99 , replace = FALSE)
  polygon_data_low_train <- tc_no[no,]
  
  # combine all random observations
  train <- do.call("rbind", list(polygon_data_high_train, polygon_data_low_train))
  
  tc <- glm(new_model_class~gv_2018, data=train, family="binomial")
  sum <-summary(tc)
  sum
  aic<- sum$aic
  aic
  model_aic <- append(model_aic, aic)
  
  # predict data 
  pred<-predict(tc, train, type="response") 
  
  # error rate for data
  pred1 <- ifelse (pred>0.5,1,0)
  tab<- table(Predicted=pred1, Actual=train$new_model_class)
  write.table(tab, file = "type conversion train response variable accuracy.csv", sep=",",col.names=NA, append=T)

  acc <-1-sum(diag(tab))/sum(tab)   #error calculation. gives % observations predicted incorrectly 
  acc <- 1- acc 
  acc_list <- append(acc_list, acc)
  
  #auc score
  pred <- predict(tc,train)
  pred1 <-prediction(pred, train$new_model_class)
  roc <- performance(pred1, "tpr", "fpr")
  auc <-performance(pred1, "auc")
  auc1 <- unlist(slot(auc, "y.values"))
  auc_list <- append(auc_list, auc1)
  
}


model_aic <- as.data.frame(unlist(model_aic))
acc_list <- as.data.frame(unlist(acc_list))
auc_list <-as.data.frame(unlist(auc_list))

combo <- cbind(model_aic, acc_list, auc_list)
write.table(model_aic, file="model_aic.csv", sep=",")

# table of mean model outputs ---------------------------------------------------

sf_combo_data <- read_csv(here::here("Shade Fraction Data Sheets and Code", "model_aic.csv"))
# the column headers are off, they 

# read in data and make it R friendly 
type_converted <- read_csv(here::here("Shade Fraction Data Sheets and Code", "type conversion train response variable accuracy.csv")) %>% 
  rename(type_converted = '0')
not_type_converted <- read_csv(here::here("Shade Fraction Data Sheets and Code", "type conversion train response variable accuracy.csv")) %>% rename(not_type_converted = '1')

type_converted <- type_converted %>% 
  select (X1, type_converted) %>% 
  drop_na() %>% 
  filter (X1 == 0) %>% 
  mutate(class_acc = type_converted/99)
type_converted_accuracy <- type_converted %>% 
   summarise(mean = mean(class_acc))

not_type_converted <- not_type_converted %>% 
  select (X1, not_type_converted) %>% 
  drop_na() %>% 
  filter (X1 == 1) %>% 
  mutate(class_acc = not_type_converted/99)
not_type_converted_accuracy <- not_type_converted %>% 
   summarise(mean = mean(class_acc))
```

## GV: two response variables 
```{r}
# make lists to later store values
acc_list <- list()
table_values <- list()
model_aic <-list()
auc_list <- list ()
#prediction_list <- list()

#type conversion model ------------------------------------
for (i in 1:500)
  {
  #divide the response variable into equal number of observations
  tc_yes <- gv2[which(gv2$new_model_class == "1"),]
  tc_no <- gv2[which(gv2$new_model_class == "0"),]
  
  # pull 175 random observations from the high class and separate
  yes <- sample(1:nrow(tc_yes), 99 , replace = FALSE)
  polygon_data_high_train <- tc_yes[yes,]
  # pull 175 random observations from the low class and separate
  no <- sample(1:nrow(tc_no), 99 , replace = FALSE)
  polygon_data_low_train <- tc_no[no,]
  
  # combine all random observations
  train <- do.call("rbind", list(polygon_data_high_train, polygon_data_low_train))
  
  tc <- glm(new_model_class~gv_2018 + gv_min_2000_2006, data=train, family="binomial")
  sum <-summary(tc)
  sum
  aic<- sum$aic
  aic
  model_aic <- append(model_aic, aic)
  
  # predict data 
  pred<-predict(tc, train, type="response") 
  
  # error rate for data
  pred1 <- ifelse (pred>0.5,1,0)
  tab<- table(Predicted=pred1, Actual=train$new_model_class)
  write.table(tab, file = "type conversion train response variable accuracy.csv", sep=",",col.names=NA, append=T)

  acc <-1-sum(diag(tab))/sum(tab)   #error calculation. gives % observations predicted incorrectly 
  acc <- 1- acc 
  acc_list <- append(acc_list, acc)
  
  #auc score
  pred <- predict(tc,train)
  pred1 <-prediction(pred, train$new_model_class)
  roc <- performance(pred1, "tpr", "fpr")
  auc <-performance(pred1, "auc")
  auc1 <- unlist(slot(auc, "y.values"))
  auc_list <- append(auc_list, auc1)
  
}


model_aic <- as.data.frame(unlist(model_aic))
acc_list <- as.data.frame(unlist(acc_list))
auc_list <-as.data.frame(unlist(auc_list))

combo <- cbind(model_aic, acc_list, auc_list)
write.table(model_aic, file="model_aic.csv", sep=",")

# table of mean model outputs ---------------------------------------------------

sf_combo_data <- read_csv(here::here("Shade Fraction Data Sheets and Code", "model_aic.csv"))
# the column headers are off, they 

# read in data and make it R friendly 
type_converted <- read_csv(here::here("Shade Fraction Data Sheets and Code", "type conversion train response variable accuracy.csv")) %>% 
  rename(type_converted = '0')
not_type_converted <- read_csv(here::here("Shade Fraction Data Sheets and Code", "type conversion train response variable accuracy.csv")) %>% rename(not_type_converted = '1')

type_converted <- type_converted %>% 
  select (X1, type_converted) %>% 
  drop_na() %>% 
  filter (X1 == 0) %>% 
  mutate(class_acc = type_converted/99)
type_converted_accuracy <- type_converted %>% 
   summarise(mean = mean(class_acc))

not_type_converted <- not_type_converted %>% 
  select (X1, not_type_converted) %>% 
  drop_na() %>% 
  filter (X1 == 1) %>% 
  mutate(class_acc = not_type_converted/99)
not_type_converted_accuracy <- not_type_converted %>% 
   summarise(mean = mean(class_acc))
```

## GV: three response variables 

```{r}
# make lists to later store values
acc_list <- list()
table_values <- list()
model_aic <-list()
auc_list <- list ()
#prediction_list <- list()

#type conversion model ------------------------------------
for (i in 1:500)
  {
  #divide the response variable into equal number of observations
  tc_yes <- gv2[which(gv2$new_model_class == "1"),]
  tc_no <- gv2[which(gv2$new_model_class == "0"),]
  
  # pull 175 random observations from the high class and separate
  yes <- sample(1:nrow(tc_yes), 99 , replace = FALSE)
  polygon_data_high_train <- tc_yes[yes,]
  # pull 175 random observations from the low class and separate
  no <- sample(1:nrow(tc_no), 99 , replace = FALSE)
  polygon_data_low_train <- tc_no[no,]
  
  # combine all random observations
  train <- do.call("rbind", list(polygon_data_high_train, polygon_data_low_train))
  
  tc <- glm(new_model_class~gv_2018 + gv_min_2000_2006 + gv_max_2000_2006, data=train, family="binomial")
  sum <-summary(tc)
  sum
  aic<- sum$aic
  aic
  model_aic <- append(model_aic, aic)
  
  # predict data 
  pred<-predict(tc, train, type="response") 
  
  # error rate for data
  pred1 <- ifelse (pred>0.5,1,0)
  tab<- table(Predicted=pred1, Actual=train$new_model_class)
  write.table(tab, file = "type conversion train response variable accuracy.csv", sep=",",col.names=NA, append=T)

  acc <-1-sum(diag(tab))/sum(tab)   #error calculation. gives % observations predicted incorrectly 
  acc <- 1- acc 
  acc_list <- append(acc_list, acc)
  
  #auc score
  pred <- predict(tc,train)
  pred1 <-prediction(pred, train$new_model_class)
  roc <- performance(pred1, "tpr", "fpr")
  auc <-performance(pred1, "auc")
  auc1 <- unlist(slot(auc, "y.values"))
  auc_list <- append(auc_list, auc1)
  
}


model_aic <- as.data.frame(unlist(model_aic))
acc_list <- as.data.frame(unlist(acc_list))
auc_list <-as.data.frame(unlist(auc_list))

combo <- cbind(model_aic, acc_list, auc_list)
write.table(model_aic, file="model_aic.csv", sep=",")

# table of mean model outputs ---------------------------------------------------

sf_combo_data <- read_csv(here::here("Shade Fraction Data Sheets and Code", "model_aic.csv"))
# the column headers are off, they 

# read in data and make it R friendly 
type_converted <- read_csv(here::here("Shade Fraction Data Sheets and Code", "type conversion train response variable accuracy.csv")) %>% 
  rename(type_converted = '0')
not_type_converted <- read_csv(here::here("Shade Fraction Data Sheets and Code", "type conversion train response variable accuracy.csv")) %>% rename(not_type_converted = '1')

type_converted <- type_converted %>% 
  select (X1, type_converted) %>% 
  drop_na() %>% 
  filter (X1 == 0) %>% 
  mutate(class_acc = type_converted/99)
type_converted_accuracy <- type_converted %>% 
   summarise(mean = mean(class_acc))

not_type_converted <- not_type_converted %>% 
  select (X1, not_type_converted) %>% 
  drop_na() %>% 
  filter (X1 == 1) %>% 
  mutate(class_acc = not_type_converted/99)
not_type_converted_accuracy <- not_type_converted %>% 
   summarise(mean = mean(class_acc))
```

## GV: four response variables 
```{r}
# make lists to later store values
acc_list <- list()
table_values <- list()
model_aic <-list()
auc_list <- list ()
#prediction_list <- list()

#type conversion model ------------------------------------
for (i in 1:500)
  {
  #divide the response variable into equal number of observations
  tc_yes <- gv2[which(gv2$new_model_class == "1"),]
  tc_no <- gv2[which(gv2$new_model_class == "0"),]
  
  # pull 175 random observations from the high class and separate
  yes <- sample(1:nrow(tc_yes), 99 , replace = FALSE)
  polygon_data_high_train <- tc_yes[yes,]
  # pull 175 random observations from the low class and separate
  no <- sample(1:nrow(tc_no), 99 , replace = FALSE)
  polygon_data_low_train <- tc_no[no,]
  
  # combine all random observations
  train <- do.call("rbind", list(polygon_data_high_train, polygon_data_low_train))
  
  tc <- glm(new_model_class~gv_2018 + gv_min_2000_2006, gv_max_2000_2006 + gv_avg_2000_2006, data=train, family="binomial")
  sum <-summary(tc)
  sum
  aic<- sum$aic
  aic
  model_aic <- append(model_aic, aic)
  
  # predict data 
  pred<-predict(tc, train, type="response") 
  
  # error rate for data
  pred1 <- ifelse (pred>0.5,1,0)
  tab<- table(Predicted=pred1, Actual=train$new_model_class)
  write.table(tab, file = "type conversion train response variable accuracy.csv", sep=",",col.names=NA, append=T)

  acc <-1-sum(diag(tab))/sum(tab)   #error calculation. gives % observations predicted incorrectly 
  acc <- 1- acc 
  acc_list <- append(acc_list, acc)
  
  #auc score
  pred <- predict(tc,train)
  pred1 <-prediction(pred, train$new_model_class)
  roc <- performance(pred1, "tpr", "fpr")
  auc <-performance(pred1, "auc")
  auc1 <- unlist(slot(auc, "y.values"))
  auc_list <- append(auc_list, auc1)
  
}


model_aic <- as.data.frame(unlist(model_aic))
acc_list <- as.data.frame(unlist(acc_list))
auc_list <-as.data.frame(unlist(auc_list))

combo <- cbind(model_aic, acc_list, auc_list)
write.table(model_aic, file="model_aic.csv", sep=",")

# table of mean model outputs ---------------------------------------------------

sf_combo_data <- read_csv(here::here("Shade Fraction Data Sheets and Code", "model_aic.csv"))
# the column headers are off, they 

# read in data and make it R friendly 
type_converted <- read_csv(here::here("Shade Fraction Data Sheets and Code", "type conversion train response variable accuracy.csv")) %>% 
  rename(type_converted = '0')
not_type_converted <- read_csv(here::here("Shade Fraction Data Sheets and Code", "type conversion train response variable accuracy.csv")) %>% rename(not_type_converted = '1')

type_converted <- type_converted %>% 
  select (X1, type_converted) %>% 
  drop_na() %>% 
  filter (X1 == 0) %>% 
  mutate(class_acc = type_converted/99)
type_converted_accuracy <- type_converted %>% 
   summarise(mean = mean(class_acc))

not_type_converted <- not_type_converted %>% 
  select (X1, not_type_converted) %>% 
  drop_na() %>% 
  filter (X1 == 1) %>% 
  mutate(class_acc = not_type_converted/99)
not_type_converted_accuracy <- not_type_converted %>% 
   summarise(mean = mean(class_acc))
```


## GV: four response variables 
```{r}
# make lists to later store values
acc_list <- list()
table_values <- list()
model_aic <-list()
auc_list <- list ()
#prediction_list <- list()

#type conversion model ------------------------------------
for (i in 1:500)
  {
  #divide the response variable into equal number of observations
  tc_yes <- gv2[which(gv2$new_model_class == "1"),]
  tc_no <- gv2[which(gv2$new_model_class == "0"),]
  
  # pull 175 random observations from the high class and separate
  yes <- sample(1:nrow(tc_yes), 99 , replace = FALSE)
  polygon_data_high_train <- tc_yes[yes,]
  # pull 175 random observations from the low class and separate
  no <- sample(1:nrow(tc_no), 99 , replace = FALSE)
  polygon_data_low_train <- tc_no[no,]
  
  # combine all random observations
  train <- do.call("rbind", list(polygon_data_high_train, polygon_data_low_train))
  
  tc <- glm(new_model_class~gv_2018 + gv_min_2000_2006, gv_max_2000_2006 + prefire, data=train, family="binomial")
  sum <-summary(tc)
  sum
  aic<- sum$aic
  aic
  model_aic <- append(model_aic, aic)
  
  # predict data 
  pred<-predict(tc, train, type="response") 
  
  # error rate for data
  pred1 <- ifelse (pred>0.5,1,0)
  tab<- table(Predicted=pred1, Actual=train$new_model_class)
  write.table(tab, file = "type conversion train response variable accuracy.csv", sep=",",col.names=NA, append=T)

  acc <-1-sum(diag(tab))/sum(tab)   #error calculation. gives % observations predicted incorrectly 
  acc <- 1- acc 
  acc_list <- append(acc_list, acc)
  
  #auc score
  pred <- predict(tc,train)
  pred1 <-prediction(pred, train$new_model_class)
  roc <- performance(pred1, "tpr", "fpr")
  auc <-performance(pred1, "auc")
  auc1 <- unlist(slot(auc, "y.values"))
  auc_list <- append(auc_list, auc1)
  
}


model_aic <- as.data.frame(unlist(model_aic))
acc_list <- as.data.frame(unlist(acc_list))
auc_list <-as.data.frame(unlist(auc_list))

combo <- cbind(model_aic, acc_list, auc_list)
write.table(model_aic, file="model_aic.csv", sep=",")

# table of mean model outputs ---------------------------------------------------

sf_combo_data <- read_csv(here::here("Shade Fraction Data Sheets and Code", "model_aic.csv"))
# the column headers are off, they 

# read in data and make it R friendly 
type_converted <- read_csv(here::here("Shade Fraction Data Sheets and Code", "type conversion train response variable accuracy.csv")) %>% 
  rename(type_converted = '0')
not_type_converted <- read_csv(here::here("Shade Fraction Data Sheets and Code", "type conversion train response variable accuracy.csv")) %>% rename(not_type_converted = '1')

type_converted <- type_converted %>% 
  select (X1, type_converted) %>% 
  drop_na() %>% 
  filter (X1 == 0) %>% 
  mutate(class_acc = type_converted/99)
type_converted_accuracy <- type_converted %>% 
   summarise(mean = mean(class_acc))

not_type_converted <- not_type_converted %>% 
  select (X1, not_type_converted) %>% 
  drop_na() %>% 
  filter (X1 == 1) %>% 
  mutate(class_acc = not_type_converted/99)
not_type_converted_accuracy <- not_type_converted %>% 
   summarise(mean = mean(class_acc))
```









